Industrial Applications of CFD

### 2011-03-10 MSTE_methodology.tex
we have equations for the numerical example now!
agregue conclusiones y escribi sobre el numerical method

\TODO{     * Compare the MSTE with other internal variable problems and suggest LSQ as a possibility.}

@@

\TODO{[REESCRIBIR ECUACIONES]}

\TODO{Mention implementation details (polynomials and quadratures used, number of points, solver, etc.).  ouch... }

\TODO{Explain and show q, epsilon and h. and BCs}

@@

\TODO{Describe what case could be described with this simplified model.}

\TODO{Show solution (phi) and S.}

\InsFig{solution}{Solution}{solution}
\InsFig{source}{Source}{source}

\TODO{Explain what is happening with phi and S in each scale, and the mixing by h.}

@@

\TODO{Contribution of this paper:}

\TODO{    * This paper proposes and explains how to solve MSTE with LSQ.}
\TODO{    * Shows that LSQ solves the MSTE with exponential convergence.}

\TODO{mas despues que tengamos todas las figuras}

>>>>

Many other phenomena in nature can be described by an integro-differential equation with one or many internal coordinates. The Boltzmann transport equation has the molecule velocity as an internal coordinate. The transport equation applied to neutrons or its condensation into the diffusion equation for neutrons use the neutron energy as internal energy. Population balance problems use the particle, bubble or droplet size as internal coordinate. The least square method has been applied succesfully to population balance problems by \cite{Dorao05a}.

@@

We select the case of flow through a column for our numerical example. A one metre column filled with a packed bed of rocks has a porosity per scale given by $\epsilon({\xi})$. Liquid is introduced from the top which percoles down the column at a given discharge per scale given by $q({\xi})$. The pore-scale redistribution, $h$  is assumed proportional to the discharges.

In the manufacturing solution method we propose a solution $\phi(\xi, z)$ and calculate analitically the value of the source. For this problem the values chosen are:

\begin{align}
\epsilon(\xi)=&A (\xi-\xi_0) (-\xi+\xi_f) \\
q({\xi}) =&V \epsilon(\xi) \xi^2\\
h(\hat \xi, \xi, z) =& q(\hat \xi) q(\xi)/L\\
\phi(\xi, z) =& 10 z (1-M Log({\xi}/R))\\
\nonumber S(\xi, z)=& 10 A V {\xi}^2 ({\xi}-\xi_0) (-{\xi}+\xi_f) \left(1-M \text{Log}\left[\frac{{\xi}}{R}\right]\right) +\\
\nonumber  &A^2 M V^2  (360 L)^{-1} {\xi}^2 ({\xi}-\xi_0) ({\xi}-\xi_f) z \times ... \\
\nonumber  &\big(-81 \xi_0^5+175 \xi_0^4 \xi_f-175 \xi_0 \xi_f^4+  ...\\
\nonumber &81 \xi_f^5+60 \xi_0^4 (3 \xi_0-5 \xi_f) \text{Log}\left[\frac{\xi_0}{{\xi}}\right]+ ...\\
& 60 (5 \xi_0-3 \xi_f) \xi_f^4 \text{Log}\left[\frac{\xi_f}{{\xi}}\right]\big)
\end{align}

\noindent with A=, V=, M= and M= \TODO{no me acuerdo}.

\TODO{Mention implementation details (polynomials and quadratures used, number of points, solver, etc.).  ouch... GLL, 12 puntos como se ve en la figura}

%\TODO{Explain and show q, epsilon and h. and BCs}

Both the discharge, $q({\xi})$ and the porosity, $\epsilon({\xi})$ do not depend on the position along the column. Plot of the distribution and cummulative distribution for these two functions are given in Figure \ref{fig:profiles}.

@@

%\TODO{Describe what case could be described with this simplified model.}

The solution is shown in Figure \ref{fig:solution}. As the liquid flows down the column $\phi$ increases due to reactions happening inside the pores. Despite the source term is very different for different pore sizes (negative for large pores and positive for small pores, see Figure \ref{fig:source}) the tracer $\phi$ is exchanged between different pores scales tending to level out the differences. However the increase is larger for the small pores at all the length of the column (z values). The source term (Figure \ref{fig:source}) behaves differently for different pore scales. Large amounts of $\phi$ are generated for pores smaller than 5 mm, while a sink is observed in the largest pores. This behaviour for the source term mimic two compiting reactions. It models the case of a specie that is a product of an heterogeneous reaction but is a reactant of a reaction taking place in the liquid bulk. Heterogeneous reactions depend strongly of the surface to volume ratio which is much larger for small pore scales. At the same time the reactant is consumed in the larger pores in this case as the bulk reaction that consumes $\phi$ becomes more dominant.

\InsFig{solution}{Solution $\phi(z,\xi)$.}{solution}

\InsFig{source}{Source term $S(z,\xi)$. }{source}

%\TODO{Explain what is happening with phi and S in each scale, and the mixing by h.}

@@

The conclusions of this paper are:

\begin{itemize}
\item The least square high order method is well suited for the solution of the multi-scale transport equation for porous media achieving exponencial convergence in the method order.
\item A method has been described and can now be applied for the solution of transport and reactions problems in porous media.
\end{itemize}

Future work still remains on the area of finding the most adequate kernels that best represent different porous media. Mainly empirical studies of the redistribution function need to be done.
